---
title: "Rarotonga shaded vs exposed Photobiology"
author: "Rob Lewis"
date: "26/04/2021"
output: html_document
---
# libraries

```{r}
library(tidyverse)
library(forcats)
library(lubridate)
library(broom)
library(nls.multstart)
library(nlstools)
#library(car)
library(pavo)
library(ggspectra)
library(photobiology)
library(photobiologyWavebands)
library(photobiologyInOut)
library(cowplot)

unnest <- unnest_legacy



```

#Load objects
```{r}
PAM <- readRDS("PAM.RDS")

FqFmparams <- readRDS("FqFmparams_no_PAR_adjust.RDS")

ETRparams <- readRDS("ETRparams.rds")

onemC_max_EnormEk <- readRDS("onemC_max_EnormEk.rds") %>%
  filter(sampleID != "C96" & sampleID != "E96" & sampleID != "C49" & sampleID != "E49")

onemQ_max_EnormEk <- readRDS("onemQ_max_EnormEk.rds") %>%
  filter(sampleID != "C96" & sampleID != "E96" & sampleID != "C49" & sampleID != "E49")

dynamQ_max_EnormEk <- readRDS("dynamQ_max_EnormEk.RDS")

total_brightness <- readRDS("total_brightness.rds")

```

# Read "PAM" excel file

```{r}
PAM <- read_csv("PAM_data/PAM.csv", col_names = TRUE) %>%
  group_by(PAMID) %>%
  mutate(date = dmy(time_date),
         Fm = ifelse(Fm <= F, F+1, Fm),
         PAR = ifelse(PAR == 0, 0.001, PAR),
         FqFm = (Fm - F) / Fm,
         rETR = FqFm * PAR,
         Fo.p = first(F) / (first(FqFm) + (first(F) / Fm)),
         onemC = (Fm - F) / (Fm - Fo.p),
         Fv.p = Fm - Fo.p,
         onemQ = (Fv.p/Fm) / first(FqFm)) %>%
  ungroup() %>%
  filter(sampleID != "C96" & sampleID != "E96" & sampleID != "C49" & sampleID != "E49")

saveRDS(PAM, "PAM.RDS")
```


# Load Hennige equation

```{r}
Hennige <- function(FqFmmax, Ek, x){
  model <- ((FqFmmax * Ek) * (1-exp(-x/Ek)))/x 
  return(model)
}
```

# Fit Fq/Fm data with Hennige equation for each replicate

```{r}
FqFmfits <- PAM %>% 
  group_by(PAMID, genus, species, habitat, replicate, sampleID) %>%
  nest() %>% 
  mutate(fit = map(data, ~ nls_multstart(FqFm ~ Hennige(FqFmmax, Ek, x = PAR),
                                     data = .x, 
                                     iter = 250, 
                                     start_lower = c(FqFmmax = 0.2, Ek = 5),
                                     start_upper = c(FqFmmax = 0.85, Ek = 1409),
                                     supp_errors = "Y",
                                     convergence_count = 100,
                                     na.action = na.omit, 
                                     lower = c(FqFmmax = 0.1, Ek = 5))))
```

```{r}
FqFminfo <- FqFmfits %>%
  unnest(fit %>% map(glance))

FqFmparams <- FqFmfits %>% 
  unnest(fit %>% map(tidy))

FqFmCI <- FqFmfits %>%
  unnest(fit %>% map(~ confint2(.x, level = 0.95) %>% data.frame() %>% dplyr::rename(conf.low = X2.5.., conf.high = X97.5..))) %>% 
  group_by(PAMID) %>%
  dplyr::mutate(term = c("FqFmmax", "Ek")) %>% 
  ungroup()

FqFmparams <- left_join(FqFmparams, FqFmCI)

new_preds <- PAM %>% 
  do(data.frame(PAR = seq(min(.$PAR), max(.$PAR),length.out = 250), stringsAsFactors = FALSE))

FqFmpreds <- FqFmfits %>% 
  unnest(fit %>% map(augment, newdata = new_preds)) %>% 
  dplyr::rename(FqFm = .fitted)

saveRDS(FqFmparams, "FqFmparams_no_PAR_adjust.RDS")
```

# Plot Fq/Fm model fits

```{r}
ggplot() + 
  geom_point(aes(PAR, FqFm, fill = habitat), shape = 21, size = 2, PAM) + 
  geom_line(aes(PAR, FqFm, group = PAMID, col = habitat, linetype = factor(replicate)), FqFmpreds) + 
  facet_wrap(~ species)
```

## (S 2.1) plot intra-colony Fq/Fmmax fold diff vs species
```{r}

FqFmmax_fold_diff <- FqFmparams %>%
  filter(term == "FqFmmax") %>%
  group_by(habitat) %>%
  dplyr::mutate(n = row_number()) %>%
  pivot_wider(names_from = habitat, values_from = c(estimate, std.error, statistic, p.value), id_cols = c(genus, species, n)) %>%
  mutate(FqFmmax_fold_change = (estimate_cryptic - estimate_exposed) / estimate_exposed) %>%
  ggplot() +
  geom_boxplot(aes(reorder(species, FqFmmax_fold_change, mean),  FqFmmax_fold_change)) +
  theme(aspect.ratio = 1) +
  ylab("Intra-colony Fq/Fm max fold difference") +
  xlab("Species") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = 1, vjust = .5, face = "italic"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"))


#ggsave("S1_FqFmmax_fold_diff.pdf", plot = FqFmmax_fold_diff, units = "in", height = 7, width = 10)
```

## (S 2.2) plot intra-colony Ek fold diff vs Species
```{r}
S2_Ek_fold_diff <- FqFmparams %>%
  filter(term == "Ek") %>%
  group_by(habitat) %>%
  dplyr::mutate(n = row_number()) %>%
  pivot_wider(names_from = habitat, values_from = c(estimate, std.error, statistic, p.value), id_cols = c(genus, species, n)) %>%
  mutate(Ek_fold_change = (estimate_exposed - estimate_cryptic) / estimate_cryptic) %>%
  ggplot() +
  geom_boxplot(aes(reorder(species, Ek_fold_change, mean),  Ek_fold_change)) +
  theme(aspect.ratio = 1) +
  ylab("Intra-colony Ek fold difference") +
  xlab("Species") +
  theme_bw() + 
  theme(panel.border = element_rect(fill = NULL, colour = "black"), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = 1, vjust = .5, face = "italic"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"))

#ggsave("S2_Ek_fold_diff.pdf", plot = S2_Ek_fold_diff, units = "in", height = 7, width = 10)
```



```{r}
# Jassby and Platt 1976 model
JP <- function(x, a, Ek){
  model <- a*Ek*tanh(x/Ek)
  return(model)
}
```

```{r}
ETRfits <- PAM %>% 
  group_by(PAMID, habitat, genus, species, replicate) %>%
  nest() %>%
  mutate(fit = map(data, ~ nls_multstart(rETR ~ JP(x = PAR, a, Ek),
                                    data = .x, 
                                     iter = 250, 
                                     start_lower = c(a = 0.05, Ek = 5),
                                     start_upper = c(a = 0.95, Ek = 1409),
                                     supp_errors = "Y",
                                     convergence_count = 100,
                                     na.action = na.omit, 
                                     lower = c(a = 0.05, Ek = 5))))
```

```{r}
ETRinfo <- ETRfits %>%
  unnest(fit %>% map(glance))

ETRparams <- ETRfits %>% 
  unnest(fit %>% map(tidy))

ETRCI <- ETRfits %>%
  unnest(fit %>% map(~ confint2(.x, level = 0.95) %>% data.frame() %>% dplyr::rename(conf.low = X2.5.., conf.high = X97.5..))) %>% 
  group_by(PAMID) %>%
  dplyr::mutate(term = c("a", "Ek")) %>% 
  ungroup()

ETRparams <- left_join(ETRparams, ETRCI)

new_preds <- PAM %>% 
  do(data.frame(PAR = seq(min(.$PAR),max(.$PAR),length.out = 250),stringsAsFactors = FALSE))

ETRpreds <- ETRfits %>% 
  unnest(fit %>% map(augment, newdata = new_preds)) %>% 
  dplyr::rename(rETR = .fitted)

saveRDS(ETRparams, "ETRparams.rds")
```

```{r}
ggplot() + 
  geom_point(aes(PAR, rETR, fill = habitat), shape = 21, size = 2, PAM) + 
  geom_line(aes(PAR, rETR, group = PAMID, col = habitat, linetype = factor(replicate)), ETRpreds) + 
  facet_wrap(~ species)

```



```{r}
ETR_Ek_fold <- ETRparams %>%
  filter(term == "Ek") %>%
  group_by(habitat) %>%
  dplyr::mutate(n = row_number()) %>%
  pivot_wider(names_from = habitat, values_from = c(estimate, std.error, statistic, p.value), id_cols = c(genus, species, n)) %>%
  mutate(Ek_fold_change = (estimate_exposed - estimate_cryptic)/estimate_cryptic, 
         species = fct_reorder(species, Ek_fold_change, .fun = "median"))

ggplot(ETR_Ek_fold) + 
  geom_boxplot(aes(species, Ek_fold_change)) +
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.1))
```

#Fq/Fm_max summary and one-way ANOVA

```{r}
FqFm_max_summary <- FqFmparams %>%
  filter(term == "FqFmmax")


library(car)

results_FqFmmax_ANOVA <- aov(estimate ~ habitat, data = FqFm_max_summary)

summary(results_FqFmmax_ANOVA)

#               Df Sum Sq Mean Sq F value Pr(>F)    
# habitat       1 0.2788 0.27880   122.1 <2e-16 ***
# Residuals   198 0.4521 0.00228 

## Homogeneity of variances

plot(results_FqFmmax_ANOVA, 1)


leveneTest(estimate ~ habitat, data = FqFm_max_summary)

#           Df F value  Pr(>F)  
#   group   1  4.9651 0.02699 *
#          198  

# variance across groups is significantly different -- non-parametric test required

kruskal.test(estimate ~ habitat, data = FqFm_max_summary)

#   Kruskal-Wallis chi-squared = 81.776, df = 1, p-value < 2.2e-16

detach("package:car", unload=TRUE)

FqFm_max_summary_plot <- FqFmparams %>%
  filter(term == "FqFmmax") %>%
  ggplot() +
  geom_boxplot(aes(habitat, estimate, fill = habitat)) +
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.1)) +
  ylab("Fq/FmMax (dimensionless)") +
  xlab("Habitat") + 
  scale_x_discrete(labels = c("shaded", "exposed")) +
  scale_fill_manual(values = c("grey", "white")) +
  theme_bw() + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(color = "black", size = 12, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        legend.position = "none")

FqFm_max_summary_plot
ggsave("FqFm_max_mean_fig1.pdf", plot = FqFm_max_summary_plot, height = 5, width = 6)

```


#Ek summary and one-way ANOVA

```{r}
Ek_summary <- ETRparams %>%
  filter(term == "Ek")

library(car)

results_Ek_ANOVA <- aov(estimate ~ habitat, data = Ek_summary)

summary(results_Ek_ANOVA)

# Homogeneity of variances
plot(results_Ek_ANOVA, 1)


leveneTest(estimate ~ habitat, data = Ek_summary)

#       Df F value Pr(>F)
#group   1  2.6605 0.1045
#      198 

# variance across groups is not significantly different  

# Normality

plot(results_Ek_ANOVA, 2)

#Shapiro-Wilk test

residuals_Ek_ANOVA <- residuals(object = results_Ek_ANOVA)

shapiro.test(x = residuals_Ek_ANOVA)

# data:  ANOVA_residuals
# W = 0.91047, p-value = 1.231e-09

# normality violated -- non-parametric test required

kruskal.test(estimate ~ habitat, data = Ek_summary)


	#   Kruskal-Wallis rank sum test

#   data:  estimate by habitat
#   Kruskal-Wallis chi-squared = 68.407, df = 1, p-value < 2.2e-16

detach("package:car", unload=TRUE)

Ek_summary_plot <- FqFmparams %>%
  filter(term == "Ek") %>%
  ggplot() +
  geom_boxplot(aes(habitat,  estimate, fill = habitat)) +
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.1)) +
  ylab("Ek (µmol photons m-2 s-1)") +
  xlab("Habitat") + 
  scale_x_discrete(labels = c("shaded", "exposed")) +
  scale_fill_manual(values = c("grey", "white")) +
  theme_bw() + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(color = "black", size = 12, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        legend.position = "none")

Ek_summary_plot
ggsave("Ek_mean_fig1.pdf", plot = Ek_summary_plot, height = 5, width = 6)

```


```{r}
ETR_Ek_fold <- ETRparams %>%
  filter(term == "Ek") %>%
  group_by(habitat) %>%
  dplyr::mutate(n = row_number()) %>%
  pivot_wider(names_from = habitat, values_from = c(estimate, std.error, statistic, p.value), id_cols = c(genus, species, n)) %>%
  mutate(Ek_fold_change = (estimate_exposed - estimate_cryptic)/estimate_cryptic)

ggplot(ETR_Ek_fold) + 
  geom_boxplot(aes(species, Ek_fold_change))
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.1))


```

```{r}
dynamicQ <- PAM %>%
  filter(PAR != 0.001)

ETREks <- ETRparams %>%
  filter(term == "Ek")

dynamicQEk <- left_join(dynamicQ, ETREks) %>%
  mutate(EnormEk = PAR/estimate)
```

```{r}
OPD <- function(q0, plat, K, x) {
  model <- ((q0 - plat) * exp(-K * x) + plat)
  return(model)
}
```

```{r}
onemCfits <- dynamicQEk %>% 
  group_by(PAMID, habitat, genus, species, replicate, sampleID) %>%
  nest() %>%
  mutate(fit = map(data, ~ nls_multstart(onemC ~ OPD(x = EnormEk, q0, plat, K),
                                    data = .x, 
                                     iter = 250, 
                                     start_lower = c(q0 = 0.1, plat = 0.01, K = 0.05),
                                     start_upper = c(q0 = 1, plat = 0.9, K = 0.8),
                                     supp_errors = "Y",
                                     convergence_count = 100,
                                     na.action = na.omit, 
                                     lower = c(q0 = 0.05, plat = 0.001, K = 0.04))))
```

```{r}
onemCinfo <- onemCfits %>%
  unnest(fit %>% map(glance))

onemCparams <- onemCfits %>% 
  unnest(fit %>% map(tidy))

onemCCI <- onemCfits %>%
  unnest(fit %>% map(~ confint2(.x, level = 0.95) %>% data.frame() %>% dplyr::rename(conf.low = X2.5.., conf.high = X97.5..))) %>% 
  group_by(PAMID) %>%
  dplyr::mutate(term = c("q0", "plat", "K")) %>% 
  ungroup()

onemCparams <- left_join(onemCparams, onemCCI)

Q_preds <- dynamicQEk %>% 
  do(data.frame(EnormEk = seq(min(.$EnormEk),max(.$EnormEk),length.out = 1000),stringsAsFactors = FALSE))

Qmax_min <- dynamicQEk %>%
  group_by(PAMID) %>%
  dplyr::summarise(min_EnormEk = min(EnormEk), max_EnormEk = max(EnormEk))

onemCpreds <- onemCfits %>% 
  unnest(fit %>% map(augment, newdata = Q_preds)) %>% 
  left_join(., Qmax_min) %>%
  group_by(PAMID) %>%
  filter(EnormEk > unique(min_EnormEk) & EnormEk < unique(max_EnormEk)) %>%
  dplyr::rename(onemC = .fitted) %>%
  ungroup()
```

```{r}
onemQfits <- dynamicQEk %>% 
  group_by(PAMID, habitat, genus, species, replicate, sampleID) %>%
  nest() %>%
  mutate(fit = map(data, ~ nls_multstart(onemQ ~ OPD(x = EnormEk, q0, plat, K),
                                    data = .x, 
                                     iter = 250, 
                                     start_lower = c(q0 = 0.1, plat = 0.01, K = 0.05),
                                     start_upper = c(q0 = 1, plat = 0.9, K = 0.8),
                                     supp_errors = "Y",
                                     convergence_count = 100,
                                     na.action = na.omit, 
                                     lower = c(q0 = 0.05, plat = 0.001, K = 0.04))))
```

```{r}
onemQinfo <- onemQfits %>%
  unnest(fit %>% map(glance))

onemQparams <- onemQfits %>% 
  unnest(fit %>% map(tidy))

onemQCI <- onemQfits %>%
  unnest(fit %>% map(~ confint2(.x, level = 0.95) %>% data.frame() %>% dplyr::rename(conf.low = X2.5.., conf.high = X97.5..))) %>% 
  group_by(PAMID) %>%
  dplyr::mutate(term = c("q0", "plat", "K")) %>% 
  ungroup()

onemQparams <- left_join(onemQparams, onemQCI)

Q_preds <- dynamicQEk %>% 
  do(data.frame(EnormEk = seq(min(.$EnormEk),max(.$EnormEk),length.out = 1000),stringsAsFactors = FALSE))

Qmax_min <- dynamicQEk %>%
  group_by(PAMID) %>%
  dplyr::summarise(min_EnormEk = min(EnormEk), max_EnormEk = max(EnormEk))

onemQpreds <- onemQfits %>% 
  unnest(fit %>% map(augment, newdata = Q_preds)) %>% 
  left_join(., Qmax_min) %>%
  group_by(PAMID) %>%
  filter(EnormEk > unique(min_EnormEk) & EnormEk < unique(max_EnormEk)) %>%
  dplyr::rename(onemQ = .fitted) %>%
  ungroup()
```

```{r}
ggplot() +
  geom_line(aes(EnormEk, onemC, group = PAMID, col = species), onemCpreds) + 
  geom_point(aes(EnormEk, onemC, col = species), dynamicQEk) + 
  geom_line(aes(EnormEk, onemQ, group = PAMID, col = species), onemQpreds) + 
  geom_point(aes(EnormEk, onemQ, col = species), shape = 17, dynamicQEk) + 
  facet_wrap(~ habitat) +
  theme(aspect.ratio = 1)

```

```{r}
onemCEEK1 <- onemCpreds %>%
  group_by(PAMID) %>%
  slice(which.min(abs(EnormEk - 1)))

onemQEEK1 <- onemQpreds %>%
  group_by(PAMID) %>%
  slice(which.min(abs(EnormEk - 1)))

dynamQEEk1 <- left_join(onemCEEK1, onemQEEK1)

dynamQEEk1 %>%
  group_by(genus, species, habitat) %>%
  dplyr::summarise(onemC_mean = mean(onemC),
            onemQ_mean = mean(onemQ)) %>%
  ggplot(aes(onemC_mean, onemQ_mean, group = species)) +
  geom_line() +
  geom_point(aes(fill = species, shape = habitat), size = 4) +
  geom_abline(intercept = 0, slope = 1) +
  scale_shape_manual(values = c(21, 22)) +
  coord_cartesian(ylim = c(0.4, 1), xlim = c(0.4, 1)) +
  guides(fill = guide_legend(override.aes = list(shape = 21)))

```
# 1-C vs 1-Q summary

```{r}
onemC_max_EnormEk <- onemCpreds %>%
  group_by(PAMID) %>%
  slice(which.max(abs(EnormEk)))

onemQ_max_EnormEk <- onemQpreds %>%
  group_by(PAMID) %>%
  slice(which.max(abs(EnormEk)))

dynamQ_max_EnormEk <- left_join(onemC_max_EnormEk, onemQ_max_EnormEk)


dynamQ_CI <- dynamQ_max_EnormEk %>%
 group_by(habitat) %>%
  mutate(onemC_sd = sd(onemC), onemQ_sd = sd(onemQ),
         onemC_me = qnorm(.975)*(onemC_sd/sqrt(98)), onemQ_me = qnorm(.975)*(onemQ_sd/sqrt(98))) %>%
  select(habitat, onemC_me, onemQ_me) %>%
  distinct()
  

dynamQ_max_EnormEk_plot <- dynamQ_max_EnormEk %>%
  group_by(habitat) %>%
  dplyr::summarise(onemC_mean = mean(onemC),
            onemQ_mean = mean(onemQ)) %>%
  left_join(dynamQ_CI, dynamQ_max_EnormEk, by = "habitat") %>%
  ggplot(aes(onemC_mean, onemQ_mean)) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(ylim = c(0.0, 1), xlim = c(0.0, .5)) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  geom_errorbar(aes(ymin = onemQ_mean - onemQ_me, ymax = onemQ_mean + onemQ_me), width = 0.01, size=.75, color="black") +
  geom_errorbarh(aes(xmin = onemC_mean - onemC_me, xmax = onemC_mean + onemC_me), height = 0.02, size=.75, color="black") +
  geom_point(aes(col = habitat), size = 3) +
  scale_colour_manual(labels = c("shaded", "exposed"), values = c("grey", "white")) + 
  theme_bw() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(color = "black", size = 12, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)) +
  ylab("[1 - Q] (mean ± 95% CI)") +
  xlab("[1 - C] (mean ± 95% CI)") +
  labs(col = "Habitat")
 
dynamQ_max_EnormEk_plot
ggsave("dynamQ_mean_fig1.pdf", plot = dynamQ_max_EnormEk_plot, height = 5, width = 7.2)


#ANOVA [1 - C]
library(car)

leveneTest(onemC ~ habitat, data = dynamQ_max_EnormEk)

#       Df F value   Pr(>F)   
#group   1  9.9831 0.001833 **
#      194

#variance across groups is not homogeneous -- non-parametric test required

kruskal.test(onemC ~ habitat, data = dynamQ_max_EnormEk)

# Kruskal-Wallis chi-squared = 68.777, df = 1, p-value < 2.2e-16

#ANOVA [1 - Q]

leveneTest(onemQ ~ habitat, data = dynamQ_max_EnormEk)

#       Df F value  Pr(>F)  
#group   1  5.5807 0.01915 *
#      194 

#variance across groups is not homogeneous -- non-parametric test required

kruskal.test(onemQ ~ habitat, data = dynamQ_max_EnormEk)

# Kruskal-Wallis chi-squared = 49.69, df = 1, p-value = 1.801e-12
detach("package:car", unload=TRUE)

saveRDS(dynamQ_max_EnormEk, "dynamQ_max_EnormEk.rds")

dynamQ_max_EnormEk %>%
  ungroup()%>%
  group_by(habitat) %>%
  summarise(onemC_mean = mean(onemC),
            onemQ_mean = mean(onemQ))
```

## (S 2.3) plot species [1-C] & [1-Q] vs habitat
```{r}
dynamQ_max_EnormEk %>%
  group_by(habitat) %>%
  dplyr::mutate(n = row_number()) %>%
  pivot_wider(names_from = habitat, values_from = c(onemC), id_cols = c(genus, species, n)) %>%
  mutate(onemC_fold_change = (cryptic / exposed)) %>%
  ggplot() +
  geom_boxplot(aes(reorder(species, onemC_fold_change, mean),  onemC_fold_change)) +
  theme(aspect.ratio = 1) +
  ylab("Intra-colony [1 - C] fold difference") +
  xlab("Species") +
  theme_bw() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = 1, vjust = .5, face = "italic"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"))


dynamQ_CI <- dynamQ_max_EnormEk %>%
 group_by(habitat) %>%
  mutate(onemC_sd = sd(onemC), onemQ_sd = sd(onemQ),
         onemC_me = qnorm(.975)*(onemC_sd/sqrt(98)), onemQ_me = qnorm(.975)*(onemQ_sd/sqrt(98))) %>%
  select(habitat, onemC_me, onemQ_me) %>%
  distinct()

cb_Palette <- c("#000000", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "springgreen", "darkblue", "#57142d", "#c35d80", "#d20d30", "#811596", "#90b325", "#f68054", "#e9fbff", "aquamarine", "#a6023f")

species_dynamQ_maximal_plot <- dynamQ_max_EnormEk %>%
  group_by(species, habitat) %>%
  dplyr::summarise(onemC_mean = mean(onemC),
            onemQ_mean = mean(onemQ)) %>%
  left_join(dynamQ_CI, dynamQ_max_EnormEk, by = "habitat") %>%
  ggplot(aes(onemC_mean, onemQ_mean, group = species)) +
  geom_line(linetype = "dotted") +
  geom_abline(intercept = 0, slope = 1) +
  coord_cartesian(ylim = c(0.0, 1), xlim = c(0.0, .5)) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  ylab("[1 - Q] mean (+/- 95% CI)") +
  xlab("[1 - C] mean (+/- 95% CI)") +
  theme_bw() +
  theme(axis.text.x = element_text(color = "black", size = 12, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)) +
  geom_errorbar(aes(ymin = onemQ_mean - onemQ_me, ymax = onemQ_mean + onemQ_me), width = 0.01, size=.4, color="black") +
  geom_errorbarh(aes(xmin = onemC_mean - onemC_me, xmax = onemC_mean + onemC_me), height = 0.02, size=.4, color="black") + 
  geom_point(aes(fill = species, shape = habitat), size = 4) +
  labs(fill = "Species", shape = "Habitat") +
  scale_fill_manual(values = cb_Palette) +
  scale_shape_manual(labels = c("Shaded", "Exposed"), values = c(21, 22))


#ggsave("S3_species_dynamQ_maximal_plot.pdf", plot = species_dynamQ_maximal_plot, units = "in", height = 7, width = 10)
```

# Reflectance spectra of coral tissues

```{r}
#Import, smooth, then covert to long format and multi-spec(mspct)

# Pavo rspec object
specs <- getspec(where = "ocean_optics_data/", ext = "txt", lim = c(400, 750), decimal = ".") %>%
procspec(opt = "smooth", span = 0.1)

# Create a long-format df for ggplot
map <- PAM %>%
  select(sampleID, PAMID, genus, species, site, habitat, replicate) %>%
  group_by(sampleID, PAMID, genus, species, site, habitat, replicate) %>%
  slice(1) %>%
  ungroup %>%
  filter(sampleID != "C96" & sampleID != "E96" & sampleID != "C49" & sampleID != "E49")

specslong <- specs %>%
  pivot_longer(names_to = "sampleID", values_to = "reflectance", -wl) %>%
 filter(sampleID != "C96" & sampleID != "E96" & sampleID != "C49" & sampleID != "E49")
  
specslong <- left_join(specslong, map)
```
# Explore the spectral qualities by habitat

```{r}

# An explanation of these parameters can be found by typing ?summary.rspec in the console
spec_summary <- summary(specs) %>%
  tibble::rownames_to_column(var = "sampleID") %>%
  pivot_longer(-sampleID, names_to = "variable") %>%
  left_join(., map) %>%
  filter(sampleID != "C96" & sampleID != "E96" & sampleID != "C49" & sampleID != "E49")


# Plot differences in spectral reflectance according to habitat

spec_summary %>%
  filter(variable == "B2") %>% # Choose a parameter of interest to summarise
  ggplot() +
  geom_boxplot(aes(x = habitat, y = value)) +
  facet_wrap(~ species, labeller = labeller(.multi_line = FALSE)) +
  theme(aspect.ratio = 1) +
  ylab("B2 Mean Brightness") +
  xlab("Habitat")

# Plot all replicates
specslong %>%
ggplot(aes(wl, reflectance, colour = habitat, group = PAMID)) +
  geom_line(aes(linetype = factor(replicate))) +
  facet_wrap(~ species) +
  scale_fill_identity() +
  scale_x_wl_continuous() +
  scale_colour_manual(values = c("black", "orange")) +
  theme(aspect.ratio = 1)

```

#Specs/total brightness summary and one-way ANOVA

```{r}
library(Rmisc)

Specs_cryptic_CI <- specslong %>%
  filter(habitat == "cryptic") %>%
  group.CI(reflectance ~ wl, ., ci = .95) %>%
  mutate(habitat = "cryptic")

Specs_exposed_CI <- specslong %>%
  filter(habitat == "exposed") %>%
  group.CI(reflectance ~ wl, ., ci = .95) %>%
  mutate(habitat = "exposed")

detach("package:Rmisc", unload=TRUE)

specslong_summary <- rbind(Specs_cryptic_CI, Specs_exposed_CI)



specs_summary_plot <- specslong_summary %>%
ggplot(aes(wl, reflectance.mean, fill = habitat, colour = habitat)) +
  geom_line() +
  #scale_fill_identity() +
  scale_x_wl_continuous() +
  geom_ribbon(aes(ymin = reflectance.lower, ymax = reflectance.upper), linetype=2, alpha=0.1) +
  scale_fill_manual(labels = c("shaded", "exposed"), values = c("black", "white")) +
  scale_colour_manual(values = c("black", "black")) +
  theme(aspect.ratio = 1) +
  theme_bw() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(color = "black", size = 12, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)) +
  ylab("Reflectance mean (± 95% CI)") +
  labs(colour = "Habitat") #+
  #guides(fill ="none")

specs_summary_plot
ggsave("Reflectance_mean_fig1.pdf", plot = specs_summary_plot, height = 5, width = 7.2)

total_brightness <- spec_summary %>% 
  filter(variable == "B1") %>%
  dplyr::rename(total_reflectance = value)


library(car)

results_total_brightness_ANOVA <- aov(total_reflectance ~ habitat, data = total_brightness)

summary(results_total_brightness_ANOVA)

#              Df    Sum Sq   Mean Sq F value Pr(>F)    
# habitat       1 7.981e+08 798136334   101.2 <2e-16 ***
# Residuals   194 1.530e+09   7885627

leveneTest(total_reflectance ~ habitat, data = total_brightness)

#        Df F value   Pr(>F)   
# group   1  10.754 0.001234 **
#       198                 

# variance across groups is significantly different -- non-parametric test required

kruskal.test(total_reflectance ~ habitat, data = total_brightness)

# Kruskal-Wallis chi-squared = 71.861, df = 1, p-value < 2.2e-16

detach("package:car", unload=TRUE)
```

# (S 2.4) plot absorbance fold diff vs species
```{r}

S4_total_brightness <- total_brightness %>%
  group_by(habitat) %>%
  dplyr::mutate(n = row_number()) %>%
  pivot_wider(names_from = "habitat", values_from = "total_reflectance", id_cols = c(genus, species, n)) %>%
  mutate(total_brightness_fold_change = (exposed - cryptic)/ cryptic) %>%
  ggplot() +
  geom_boxplot(aes(reorder(species, total_brightness_fold_change, mean),  total_brightness_fold_change)) +
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.1)) +
  ylab("Intra-colony total brightness fold difference") +
  xlab("Species") +
  theme_bw() + 
  theme(panel.border = element_rect(fill = NULL, colour = "black"), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = 1, vjust = .5, face = "italic"),
        axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 14, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 14, angle = 90, hjust = .5, vjust = .5, face = "plain"))

#ggsave("S4_total_brightness.pdf", plot = S4_total_brightness, units = "in", height = 7, width = 10)

```



# Combine Figure 1

```{r}
library(patchwork)

Fig_1 <- plot_grid(FqFm_max_summary_plot, Ek_summary_plot, dynamQ_max_EnormEk_plot, specs_summary_plot, ncol = 2, align = "h", axis = "bt") +  draw_plot_label(c("A", "B", "C", "D"), 
                   size = 14, 
                   c(0, 0.5, 0, 0.5),
                   c(1, 1, 0.5, 0.5))

Fig_1

ggsave("Fig_1.pdf", plot = Fig_1, units = "in", height = 8, width = 12)
```

#save RDS
```{r}
saveRDS(FqFmparams, "FqFmparams_no_PAR_adjust.rds")

saveRDS(onemC_max_EnormEk, "onemC_max_EnormEk.rds")

saveRDS(onemQ_max_EnormEk, "onemQ_max_EnormEk.rds")

saveRDS(dynamQ_max_EnormEk, "dynamQ_max_EnormEk.rds")

saveRDS(total_brightness, "total_brightness.rds")
```

